---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
require(tidyverse)
library(tidyverse)
library(tidyr)
require(tidyr)
library(knitr)
library(markdown)
```

## Loading and preprocessing the data


```{r load}
activity<-read.csv(file="C:/Users/subash/Desktop/activity.csv",header=TRUE,sep=",",colClasses=c("integer", "Date", "integer"))


```
#1.What is mean total number of steps taken per day?
```{r sum}
sum_steps<-tapply(activity$steps,activity$date,FUN=sum,na.rm=TRUE)
names(sum_steps)<-paste0("x",1:24)

df<-as.data.frame(sum_steps)
sum_hist<-gather(df,day,steps)
```

## Make a histogram of the total number of steps taken each day

```{r hist_sum, echo=TRUE}

ggplot(sum_hist,aes(steps))+geom_histogram(breaks=seq(0,22000,by=2000),col="red",fill="red")+
  labs(title="Histogram of the total number of steps taken each day") +
  labs(x="Number of steps taken each day")
```
# Calculate and report the mean and median total number of steps taken per day
```{r mean_median,echo=TRUE}
mean(sum_hist$steps,na.rm=TRUE)
median(sum_hist$steps,na.rm=TRUE)
```
#Make a time series plot
```{r time_series, echo=TRUE}
avg_steps<-tapply(activity$steps,activity$interval,FUN=mean,na.rm=TRUE)
interval_steps <- aggregate(steps ~ interval, activity, mean,na.rm=TRUE)

ggplot(interval_steps,aes(interval,steps))+geom_line()+
  labs(title="Average daily Activity pattern per 5 min interval",x="Interval",y="Average Number of days")
```
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max_avg,echo=TRUE}
interval_steps[which.max(interval_steps$steps),]
```
Maximum number of steps along day is at 8:35

#Calculate and report the total number of missing values in the dataset
```{r missing,echo=TRUE}
miss<-sum(is.na(activity$steps))
miss
```
Total number of missing values is 2304.

#Devise a strategy for filling in all of the missing values in the dataset
Fill in the missing value with average for that interval
```{r impute,echo=TRUE}
impute <- transform(activity, steps=ifelse(is.na(steps), avg_steps, steps))
total_impsteps <- tapply(impute$steps, impute$date, sum, na.rm=T)
```
#Mean and median with the imputed data
```{r imp,echo=TRUE}
impstep_mean <- mean(total_impsteps)
impstep_mean

impstep_median <- median(total_impsteps)
impstep_median
```
#Make a histogram of the total number of steps taken each day
```{r hist_imp,echo=TRUE}
hist(total_impsteps, breaks=11, 
     xlab="number of steps per day", 
     sub="(missing values imputed)",
     main="Histogram of total steps per day")
abline(v=impstep_mean, col="red", lwd=3)
abline(v=impstep_median, col="blue", lwd=3, lty=2)
legend(x="topright", legend=c("mean","median"), col=c("red","blue"), bty="n", lwd=3)
```
#Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r new,echo=TRUE}
week <- factor(weekdays(impute$date) %in% c("Saturday","Sunday"), 
               labels=c("weekday","weekend"), ordered=FALSE)

impsteps <- aggregate(impute$steps, by=list(interval=impute$interval, weekday=week), mean)
```
#Make a panel plot containing a time series plot
```{r panel,echo=TRUE}
g <- ggplot(impsteps, aes(interval/60, x))
g + geom_line() + facet_grid(weekday ~ .) +
  scale_x_continuous(breaks=0:6*4, labels=paste(0:6*4,":00", sep="")) +
  theme_bw() +
  labs(y="average number of steps in 5-min interval") +
  labs(x="time of day (h)") +
  labs(title="Daily activity pattern")
```


  